\section{Discussions and Future Works}
\subsection{Analysis of Results}

The experimental results demonstrate that the proposed cascaded RT-DETR and PARSeq system substantially outperforms the evaluated baselines on Vietnamese document OCR. The detection results reveal that EAST, despite being a strong general-purpose text detector, fails to generalize to Vietnamese document layouts without domain-specific adaptation. The F1-score improvement from 0.134 to 0.501 at IoU 0.5 highlights the value of fine-tuning on Vietnamese-specific synthetic data, and the disproportionate improvement at IoU 0.75 confirms that the GWD loss and sine-cosine angle encoding contribute meaningfully to the tightness of localization, which is critical for minimizing background noise in the crops passed to the recognition stage.

The recognition results confirm that domain-specific fine-tuning of PARSeq on Vietnamese synthetic data produces a model that surpasses all evaluated baselines, including VietOCR, which was specifically designed for Vietnamese text. The advantage of PARSeq over VietOCR is modest in CER terms but more pronounced in WER, suggesting that PARSeq's permutation-based bidirectional decoder provides marginal but consistent advantages in resolving whole-word diacritical ambiguities relative to unidirectional recognizers.

The remaining WER of 34.71\% across all evaluated systems indicates that Vietnamese text line recognition remains a challenging task, particularly for composite characters with rare tone-diacritic combinations. Analysis of recognition errors suggests that the majority of residual mistakes occur on visually similar character pairs such as \textit{ọ}/\textit{ỏ}, \textit{ấ}/\textit{ầ}, and \textit{ứ}/\textit{ừ}, where the distinguishing diacritical mark occupies only a few pixels in the rendered image. This suggests that further improvements in recognition accuracy will require either higher-resolution input processing or more sophisticated visual feature extraction for diacritical marks.

\subsection{Limitations of the Current System}

Several limitations of the current system warrant acknowledgment. First, the detection stage was trained exclusively on synthetic document images, and while the augmentation pipeline introduces substantial photometric and geometric variation, a domain gap relative to real scanned Vietnamese documents remains. Real documents exhibit ink bleeding, page yellowing, scanning artifacts, and non-uniform illumination that are difficult to fully replicate in synthetic data. The detection F1 of 0.501 at IoU 0.5 reflects this gap and is expected to improve with the inclusion of even a modest quantity of real annotated document images in the training set.

Second, the cascaded architecture is inherently susceptible to error propagation, as discussed in Section \ref{sec:error_propagation_analysis_across_stages}. Recognition accuracy is bounded by detection quality, and any detection error that produces a misaligned or truncated crop cannot be corrected at the recognition stage. The angular localization quality of the current detection model, as evidenced by the relatively low mAP@0.5:0.95 of 0.147, suggests that there is substantial room for improvement in tight bounding box fitting, particularly for near-horizontal text lines where the angle regression head must distinguish small angular deviations from zero.

Third, the system is evaluated solely on printed text and does not address handwritten Vietnamese, which presents additional challenges arising from variable stroke widths, unconstrained character forms, and the absence of consistent inter-character spacing. The Vietnamese administrative document digitization use case increasingly involves mixed printed and handwritten content, and a production-grade system would need to handle both modalities.

Fourth, the recognition model was trained on a corpus that, while large at 2,000,000 samples, is synthetically generated and may not capture the full distribution of font styles, sizes, and print quality encountered in real-world deployment. In particular, historical documents with archaic typefaces and degraded print conditions remain outside the scope of the current training distribution.

\subsection{Directions for Future Work}

Several concrete directions for future improvement follow from the identified limitations.

\textbf{Real data augmentation.} The most impactful near-term improvement would be the collection and annotation of even a small number of real Vietnamese document images to supplement the synthetic training corpus. Transfer learning experiments have consistently shown that fine-tuning on a small real dataset after pretraining on synthetic data can close a large portion of the domain gap \cite{Baek2019Synth}. A targeted annotation effort focusing on administrative document types most commonly encountered in deployment would provide high leverage.

\textbf{Improved angle regression.} The angle head architecture could be extended to predict a full probability distribution over orientations rather than a single point estimate, enabling uncertainty-aware NMS and providing the recognition stage with information about the reliability of the orientation estimate. Alternatively, the OBB representation could be replaced with a polygon-based representation that more flexibly accommodates non-rectangular text regions such as curved or perspective-distorted text lines.

\textbf{Joint fine-tuning of cascaded stages.} While the modular cascaded design offers practical advantages in terms of independent component development, the inability to propagate gradients from the recognition stage back to the detection stage precludes joint optimization. A relaxation of this constraint through differentiable cropping operations, as explored in end-to-end trainable two-stage pipelines \cite{Qiao2021}, could enable the detector to learn bounding box parameters that are specifically optimized for the downstream recognizer's sensitivity characteristics.

\textbf{Extension to handwritten and historical Vietnamese text.} Handling handwritten Vietnamese would require training data covering unconstrained script forms and the collection of annotated handwritten corpora, which are currently unavailable at scale. Adaptation of the recognition model to historical Vietnamese documents, which may contain archaic fonts, Chu Nom characters, and mixed scripts, represents a longer-term research direction of significant archival value.

\textbf{Multi-language extension.} Although currently specialized for Vietnamese, the framework is language-agnostic. Extending it to Southeast Asian languages with complex diacritics like Thai, Lao, or Khmer requires only language-specific corpora, without fundamental architectural changes. This represents a natural direction for broadening the system's impact.
