\chapter*{Acknowledgements}

\addcontentsline{toc}{chapter}{Acknowledgements}

I wish to convey my profound appreciation to my esteemed mentor, Assoc. Prof. Dr. Nguyen Thi Kim Son, whose unwavering assistance and sagacious counsel have been instrumental throughout my scholarly pursuits. Her tutelage has markedly enriched my intellectual growth, and I am at a loss to adequately express the depth of my gratitude for his encouragement and erudition. The privilege of working under her aegis at the School of Information \& Communications Technology - Hanoi University of Industry has been an extraordinary experience. I am deeply honored to have benefited from her tutelage and will forever treasure the wisdom and personal development I have attained under her assiduous guidance.

I am also immensely grateful to the School of Information \& Communications Technology, HaUI, for fostering an environment that is both nurturing and intellectually stimulating, thereby enabling students to flourish. The institution's commitment to equity and professionalism in its pedagogical ethos has been a cornerstone of my educational journey. I extend my heartfelt thanks to all the educators and instructors whose profound knowledge and unwavering dedication have made an indelible impact. Their fervor for pedagogy and readiness to mentor students have been pivotal in enhancing my comprehension, as well as that of my peers, of the intricacies of computer science.

I wish to express my sincere and heartfelt gratitude to my cherished friends for their unwavering and heartfelt companionship and support. During dark and trying moments of adversity and despondency, they have stood as an unyielding and steadfast bulwark, offering not merely their presence but also deeply empathetic and attentive listening, profound and genuine understanding, and wise and thoughtful advice. Their encouragement has shone as a radiant and uplifting beacon of hope, enabling me to persevere through the most daunting and challenging of times.

Finally, to my beloved and treasured family, I extend my deepest and most fervent gratitude for your infinite and unconditional love, steadfast and unwavering support, and the countless and selfless sacrifices you have made on my behalf. Your unshakable and inspiring faith in my abilities has imbued my endeavors with profound and enduring meaning and has served as the powerful and relentless driving force behind my every hard-earned achievement. Your presence has been a constant and unwavering source of fortitude, wisdom, and gentle guidance, and I am profoundly and eternally thankful for the absolutely indispensable and pivotal role you have played in my extraordinary and transformative academic odyssey.
