% \section{Objectives, Outline and Scope}

\section{Research Objectives}

This thesis pursues three concrete and measurable research objectives, each targeting a distinct bottleneck in the Vietnamese document OCR pipeline.

The first objective is to develop an oriented text line detector capable of accurately localizing Vietnamese text lines across a wide range of document layouts and text orientations, with tight bounding box fitting sufficient to minimize background noise and character truncation in the crops delivered to the recognition stage. This objective is operationalized through the extension of RT-DETR V2 with a sine-cosine angle regression head and GWD loss, and through the development of a synthetic detection training corpus with uniform angular coverage.

The second objective is to develop a Vietnamese text recognition model that achieves competitive character and word accuracy across the full inventory of Vietnamese composite characters, including rare tone-diacritic combinations that are systematically underrepresented in general-purpose scene text recognition training data. This objective is operationalized through vocabulary extension, class-weighted training, phonological validity constraints, and fine-tuning on a large-scale Vietnamese-specific synthetic recognition corpus.

The third objective is to integrate the detection and recognition components into a unified, modular end-to-end pipeline with a perspective-corrected inter-stage interface, and to evaluate the integrated system against established detection and recognition baselines on held-out Vietnamese document test sets. This objective provides the empirical validation that grounds the architectural and methodological claims of the thesis.

\section{Scope}

The scope of this work is delimited along three dimensions. With respect to text modality, the system is designed and evaluated exclusively for printed Vietnamese text in formal and semi-formal document contexts, including administrative correspondence, financial records, receipts, and identity documents. Handwritten Vietnamese text falls outside the scope of the current system, as it presents additional challenges from variable stroke widths and unconstrained character forms that are not addressed by the training data or model architectures adopted here. With respect to language, the system is optimized specifically for Vietnamese and is evaluated only on Vietnamese text, though the cascaded architecture is language-agnostic and could in principle be extended to other languages with complex diacritical systems through corpus and vocabulary substitution. With respect to evaluation data, the detection and recognition components are each evaluated on held-out synthetic test sets derived from the same generation pipelines used for training, and the characterization of any residual domain gap relative to real scanned documents is based on qualitative analysis and is identified as a primary direction for future work.

\section{Content}

The digitization of Vietnamese documents has become an urgent priority for administrative agencies and enterprises. Optical character recognition (OCR) serves as the key enabling technology, yet Vietnamese orthographic complexity poses substantial challenges, including a vast character inventory, minimal visual distinctions between diacritically similar characters, and diverse layouts with arbitrary text orientations.

This bachelor thesis proposes a cascaded deep learning architecture of two independently optimized stages. The first employs a fine-tuned RT-DETR V2 detector with an auxiliary angle regression head using sine-cosine parameterization and Gaussian Wasserstein Distance loss. The second uses a fine-tuned PARSeq Vision Transformer adapted to Vietnamese through vocabulary extension to approximately 310 composite tokens, class-weighted cross-entropy loss, and a phonological validity mask at decoding. A perspective-corrected inter-stage interface converts oriented text regions into upright crops, linking angular localization quality to diacritical recognition accuracy.

To address data scarcity, two independent synthetic generation pipelines are developed: a detection pipeline producing approximately 130,000 annotated document images across seven layout archetypes with full angular coverage from $0^\circ$ to $360^\circ$, and a recognition pipeline producing approximately 2,000,000 labeled crops across clean printed, augmented scene text, and mixed-background conditions.

Experiments show the system substantially outperforms established baselines. RT-DETR achieves F1@0.50 of 0.501, versus 0.134 for EAST. PARSeq achieves a Character Error Rate of 10.18\% and Word Error Rate of 34.71\%, outperforming VietOCR, PaddleOCR, Tesseract, and EasyOCR. The system is deployed as a modular pipeline with component encapsulation facilitating future upgrades without full retraining.

\section{Bachelor Thesis Outline}

To systematically address the aforementioned objectives, this thesis is organized into five chapters as follows.

\textbf{General introduction} establishes the motivation for the work, surveys the demand for Vietnamese document digitization, characterizes the technical challenges that distinguish Vietnamese OCR from the general case, and states the research objectives, scope, and content of the proposed system.

\textbf{Chapter 1: Overview of text recognition problem} surveys two areas of prior research that inform the proposed system: traditional OCR methods and their historical development, limitations, and transition toward data-driven approaches; and deep learning approaches to text detection and recognition, covering CNN-based detectors such as EAST, CRAFT, and DBNet, transformer-based detection with RT-DETR, sequence-to-sequence recognizers culminating in PARSeq, oriented bounding box detection with rotation-aware loss functions, and Vietnamese-specific adaptations including vocabulary extension and phonological validity constraints.

\textbf{Chapter 2: Methodology and implementation} presents the complete technical design of the proposed system, covering the cascaded two-stage architecture and end-to-end pipeline design with its perspective-corrected inter-stage interface, the synthetic data generation pipelines for both the detection stage (approximately 130,000 annotated images across seven layout archetypes with corrected angular coverage) and the recognition stage (approximately 2,000,000 labeled crops across three generation branches), the model optimization strategies including the sine-cosine angle head design and GWD loss configuration for RT-DETR and the vocabulary extension, class-weighted training, and phonological validity mask for PARSeq, and the system implementation covering software frameworks, model integration, and the command-line inference pipeline.

\textbf{Chapter 3: Experiments and results} reports the experimental evaluation on the MC-OCR Vietnamese receipt dataset, including the experimental setup, evaluation metrics, and the selected detection and recognition baselines, followed by quantitative results comparing RT-DETR against EAST on text line detection and PARSeq against VietOCR, PaddleOCR, Tesseract, and EasyOCR on text line recognition, with qualitative analysis of the observed performance differences.

\textbf{Conclusions and future works} summarizes the four principal technical contributions of the thesis, draws conclusions from the experimental findings confirming the effectiveness of the proposed oriented detection and Vietnamese-adapted recognition components, and outlines five concrete directions for future improvement including real data augmentation, improved angle regression, joint end-to-end fine-tuning, extension to handwritten and historical text, and multi-language generalization.
